Water desalination using polyelectrolyte hydrogel. Gibbs ensemble modelling

### 2022-07-04 gibbs.tex
changed n to N

\begin{figure*}[t]
\centering 
\subfloat[Open system \label{fig: open}]
{\includegraphics[width=0.5\textwidth]{figures/gc}}
\subfloat[Closed system \label{fig: closed}]
{\includegraphics[width=0.5\textwidth]{figures/gibbs}}
\caption{Diamond-like network in the simulation box. Color code represents
the individual ion types (red: $\na$, blue: $\cl$, yellow: $\ca$)
and the hydrogel (gray: neutral segment ($\AH$), cyan: charged segment
($\A$)). \label{fig:open and closed}.}
\end{figure*}
We simulate the compression of the gel in two ensembles.
\begin{enumerate}
\item The first is so called \emph{open system}, when the simulation box
with gel freely exchanges ions with a big amount of aqueous solution
of certain constant salinity. 
\item And in \emph{closed system}, when the gel is in equilibrium with a
finite volume of aqueous solution.
\end{enumerate}
The compression of the gel in the \emph{open} system does not affect the surrounding salinity, thus the density of ions outside the gel remains constant $\cs =$ $\cna=$ $\ccl=$ $\text{Const}$ (\reffig{fig: open}). 
On the contrary, what remains constant in the \emph{closed} system, is the total number of $\na$ and $\cl$ ions, that are contained in the gel and in the external volume, \ie in the volume $\Vbox$ (\reffig{fig: closed}).

Let $\nna$ and $\ncl$ be the density of the corresponding ion in the compression volume, $\Vbox$ averaged over this volume
\begin{equation}
n_i = \left(N_i\gel + \cs\cdot(\Vbox - \Vgel)\right) / \Vbox
\end{equation}
where $ i \in \{\na, \cl\}$,  $N_i$ is the amount of the corresponding ions in gel.
Then in closed system $\nna$ and $\ncl$ remain constant, whereas the salinity of supernate, $\cs$, may change with the change of the gel volume.

@@

\paragraph{Closed system.}
The similar scheme we use to simulate the gel in closed system. 
The procedure represents a chain of subsequent MD and MC simulation, but now the molecular dynamics is runing in two simulation boxes simultaneously.
The first box of the volume $\Vgel$ contains the gel with ions, and the second box, $\Vout$--- contains only ions. 
The total volume of both boxes is kept constant, $\Vbox=V_{\mathrm{gel}}+V_{\mathrm{out}}$, so the compression of the gel implies the decrease $\Vgel$ and increase $\Vout$.
These two boxes are represented in \reffig{fig:diamond}.

@@

The dependencies of $\Pgel$ on the gel molar volume $\Vgel$ are present in Figure \ref{fig: PV} for a set of different salinities. 
These dependencies (the case of \emph{open system}) are represented by solid lines. 
For example, the blue solid line illustrates the compression (or swelling) of the gel in equilibrium with a reservoir of salinity, $\cs=0.063$ mol/l. 
The points where the pressure equals zero, $\Pgel=0$, (indicated by filled circles) are the gel \emph{free swelling equilibrium} states. 
These states shift towards smaller volumes with increase of salinity. 
It is seen, that the increase of salinity shifts the free swelling equilibrium towards smaller volumes. 
In general, the increase of salinity shift all the $\Pgel(V)$ curves towards smaller volumes

This effect is well known and is typical for all branched \emph{strong} polyelectrolytes. 
It is caused by the decrease of ions osmotic pressure and by a screening of electrostatic interactions \cite{Zhulina2000, Landsgesel2020a}.\footnote{The salinity dependence of the size of \emph{weak} polyelectrolyte gel is in general non-monotonic. We discuss this case in \cite{Rud2018}. } %

@@

\paragraph{Compression in closed system.}
Each set of simulations in an \emph{open system} defines the free swelling equilibrium states of the gel for each salinity, that is the state where the gel partial pressure equals zero. 
These states are characterized by the corresponding molar volume of the gel, $\Vgel^0$ and the densities of ions in gel \{$\nna^0$, $\ncl^0$\} 
(Index ``0'' stands for zero bar applied pressure).
We use the free swelling equilibrium states as the starting ones for the gel compression in \emph{closed system}. 
Thus, we simulate the compression of the gel in the closed volume, $\Vbox$, which is equal to $\Vgel^0$, and contains $\ncl^0$ of $\cl$ ions and the $\na$ ions in the amount neutralizing the system,$\ncl^0$.

@@

The process of the gel compression in the \emph{closed} system is depicted in Figure~\ref{fig: PV} (dotted lines) as the dependence of gel partial pressure versus the gel molar volume.
In this plot, the blue dotted line, for example, illustrates the compression of the gel equilibrated with the solution of salinity $\cs=0.063$ mol/l, in a volume which gel has at zero pressure. 
Since, in this case, the volumes of the gel $\Vgel$ and of the supernate $\Vout$ are comparable, the compression of the gel affects the salinity of supernate, $\cs$. 
The supernate salinity now decreases, changing from $\cs^{0}=0.063$ at the free swelling equilibrium state to $\cs^{5}=0.045$ mol/l in the state with applied pressure $P\gel=5$ bar (index ``5'' stands for 5 bar).

The \reffig{fig: CV} illustrates the same swelling/compression processes, but in different coordinates: \ie salinity of supernate versus the gel molar volume, $\cs(\Vgel)$.
In these coordinates, all the open system compressions show up as horizontal lines, which reflects the constant salinity, whereas the compressions in the closed system demonstrate the decrease of $\cs$.

Although the salinity during compression in \emph{open} system remains constant, the number of ions in a compression volume, i.e. in the volume where the gel is compressed (or swells), changes.
As the compression volume $\Vbox$ we chose the volume of the free swelling equilibrium state of the gel, $\Vbox = \Vgel^0$.
In the Figure \ref{fig: NV} we depicted the number of $\cl$ ions in the volume $\Vbox$, calculated per unit volume, $\ncl =$ $\Ncl/\Vbox$, as a function of the gel molar volume. 
The depicted value can be considered as the density of $\cl$ ions averaged out over the volume $\Vbox$.

>>>>

\begin{figure*}[t]
\centering 
\subfloat[Open system \label{fig: open}]
{\includegraphics[width=0.5\textwidth]{figures/gc}}
\subfloat[Closed system \label{fig: closed}]
{\includegraphics[width=0.5\textwidth]{figures/gibbs}}
\caption{Diamond-like network in the simulation box. Color code represents
the individual ion types (red: $\na$, blue: $\cl$, yellow: $\ca$)
and the hydrogel (gray: neutral segment ($\AH$), cyan: charged segment
($\A$)). \label{fig:open and closed}.}
\end{figure*}
We simulate the compression of the gel in two ensembles.
\begin{enumerate}
\item The first is so called \emph{open system}, when the simulation box
with gel freely exchanges ions with a big amount of aqueous solution
of certain constant salinity, $\cs$. 
\item And in \emph{closed system}, when the gel is in equilibrium with a
finite volume of aqueous solution.
\end{enumerate}
The compression of the gel in the \emph{open} system does not affect the surrounding salinity, thus the density of ions outside the gel remains constant $\cs =$ $\cna=$ $\ccl=$ $\text{Const}$ (\reffig{fig: open}). 
On the contrary, what remains constant in the \emph{closed} system, is the total number of $\na$ and $\cl$ ions, that are contained in the gel and in the external volume, \ie in the volume $\Vbox$ (\reffig{fig: closed}).

%Let $\nna$ and $\ncl$ be the density of the corresponding ion in the compression volume, $\Vbox$ averaged over this volume
\begin{equation}
N_i = N_i\gel + \cs\cdot(\Vbox - \Vgel)
\end{equation}
where $i \in \{\na, \cl\}$,  $N_i$ is the amount of the corresponding ions in gel,

%Then in closed system $\nna$ and $\ncl$ remain constant, whereas the salinity of supernate, $\cs$, may change with the change of the gel volume.

@@

\paragraph{Closed system.}
The similar scheme we use to simulate the gel in closed system. 
The procedure represents a chain of subsequent MD and MC simulation, but now the molecular dynamics is runing in two simulation boxes simultaneously.
The first box of the volume $\Vgel$ contains the gel with ions, and the second box, $\Vout$--- contains only ions. 
The total volume of both boxes is kept constant, $\Vbox=V_{\mathrm{gel}}+V_{\mathrm{out}}$, so the compression of the gel implies the decreasing $\Vgel$ and increasing $\Vout$.
These two boxes are represented in \reffig{fig:diamond}.

@@

The dependencies of $\Pgel$ on the gel molar volume $\Vgel$ are present in Figure \ref{fig: PV} for a set of different salinities. 
These dependencies (the case of \emph{open system}) are represented by solid lines. 
For example, the blue solid line illustrates the compression (or swelling) of the gel in equilibrium with a reservoir of salinity, $\cs=0.063$ mol/l. 
The points where the pressure equals zero, $\Pgel=0$, (indicated by filled circles) are the gel \emph{free swelling equilibrium} states. 
These states shift towards smaller volumes with increase of salinity. 
In general, the increase of salinity shift all the $\Pgel(V)$ curves towards smaller volumes.
This effect is well known and is typical for all branched \emph{strong} polyelectrolytes. 
It is caused by the decrease of ions osmotic pressure and by a screening of electrostatic interactions \cite{Zhulina2000, Landsgesel2020a}.
\footnote{The salinity dependence of the size of \emph{weak} polyelectrolyte gel is in general non-monotonic. We discuss this case in \cite{Rud2018}.}

@@

\paragraph{Compression in closed system.}
Each set of simulations in \emph{open system} defines the free swelling equilibrium states of the gel for each salinity, \ie the state where the gel partial pressure equals zero. 
These states are characterized by the corresponding molar volume of the gel, $\Vgel^0$ and the amount of ions in gel \{$\Nna^0$, $\Ncl^0$\} 
(Index ``0'' stands for zero bar applied pressure).
We use the free swelling equilibrium states as the starting ones for the gel compression in \emph{closed system}. 
Thus, we simulate the compression of the gel in a closed volume, $\Vbox$, which is equal to $\Vgel^0$, and contains $\Ncl^0$ of $\cl$ ions and the $\na$ ions in the amount neutralizing the system, $\Nna^0 = $$\Ncl^0 + \Ngel$.

@@

The processes of the gel compression in \emph{closed} system are depicted in Figure~\ref{fig: PV} as dotted lines.
In this plot, the blue dotted line, for example, illustrates the compression of the gel equilibrated with the solution of salinity $\cs=0.063$ mol/l, in a volume which gel has at zero pressure. 
Since, in this case, the volumes of the gel $\Vgel$ and of the supernate $\Vout$ are comparable, the compression of the gel affects the salinity of supernate, $\cs$. 
The supernate salinity now decreases with compression.
This decrease is illustrated \reffig{fig: CV}, where the same swelling/compression processes are displayed in different coordinates: \ie salinity of supernate versus the gel molar volume, $\cs(\Vgel)$.
In these coordinates, all the open system compressions show up as horizontal lines, which reflects the constant salinity, whereas the compressions in closed system demonstrate the change of $\cs$ from $\cs^{0}=0.063$ at zero pressure $\Pgel$ to $\cs^{5}=0.045$ mol/l at $P\gel=5$ bar (index ``5'' stands for 5 bar).

Although the salinity during compression in \emph{open} system remains constant, the number of ions in a compression volume, \ie in the volume where the gel is compressed (or swells), changes.
%As the compression volume $\Vbox$ we chose the volume of the free swelling equilibrium state of the gel, $\Vbox = \Vgel^0$.
In the Figure \ref{fig: NV} we depicted the number of $\cl$ ions in the volume $\Vbox$, calculated per unit volume, $\ncl =$ $\Ncl/\Vbox$, as a function of the gel molar volume. 
The depicted value can be considered as the density of $\cl$ ions averaged out over the volume $\Vbox$.
